---
title: "Network generation"
author:
- name: Arsenij Ustjanzew (arsenij.ustjanzew@uni-mainz.de)<br><a href='https://www.unimedizin-mainz.de/imbei/'>IMBEI,
    University Medical Center Mainz</a><br>
date: "28.08.2023"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: show
    code_download: true
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#",
  error = FALSE,
  warning = FALSE,
  message = FALSE
)
options(width = 100)
```

```{r echo=TRUE, warning=FALSE, message=FALSE, results='hide'}
# libraries
library(NetPathMiner)
library(org.Hs.eg.db)
library(igraph)
library(ggraph)
library(graphlayouts)
library(SummarizedExperiment)
library(DESeq2)
library(DT)
library(KEGGREST)
library(rlang)
library(visNetwork)
library(pheatmap)
```

## Build metabolic network


- "hsa00600" Sphingolipid metabolism
- "hsa00601" Glycosphingolipid biosynthesis - lacto and neolacto series
- "hsa00603" Glycosphingolipid biosynthesis - globo and isoglobo series
- "hsa00604" Glycosphingolipid biosynthesis - ganglio series

```{r echo=TRUE, warning=FALSE, message=FALSE}
pathways <- c("hsa00600", "hsa00601", "hsa00603", "hsa00604")
kgml_dir <- "./data/KGML/"

if(!dir.exists(kgml_dir)) dir.create(kgml_dir)

sapply(pathways, function(p){
  kgml <- KEGGREST::keggGet(p, "kgml")
  cat(kgml, file=paste0("./data/KGML/",p,".xml"),sep="\n",append=FALSE)
})

graph <- NetPathMiner::KGML2igraph(filename = kgml_dir)
```

Recurrent metabolites (H2o, CO2 etc.) are excluded.
Disconnected nodes (metabolites, that are not connected through a reaction edge with the network) are excluded.
We get a bipartite graph containing metabolites and reactions as nodes:

**Number of verices:**
```{r echo=FALSE}
# number of verices
length(V(graph))
```

**Number of edges:**
```{r echo=FALSE}
# number of edges
length(E(graph))
```

```{r echo=FALSE}
type <- rep("m", length(V(graph)))
type[V(graph) %in% V(graph)[ reactions ]] <- "r"

ggraph(graph,layout = "stress")+
  geom_edge_link(edge_colour = "grey66",edge_width = 0.5, 
                  arrow = arrow(type = "closed",
                                angle = 25,
                                length = unit(1.5, 'mm')),
                  end_cap = circle(1, 'mm'))+
  geom_node_point(aes(fill=factor(type)), shape = 21 ,size = 3)+
  geom_node_text(aes(label = name, color=type),family = "serif",repel = TRUE, show.legend = FALSE)+
  scale_color_brewer(palette="Set1")+
  scale_fill_discrete(name="Type of edges", labels=c("Metabolite","Reaction"))+
  theme_graph()+
  theme(legend.position = "bottom")
```

The following step removes metabolite nodes keeping them as edge attributes. The resulting network contains reaction nodes only, where edges indicate that a metabolite produced by one reaction is consumed by the other.

```{r echo=TRUE}
mgraph <- makeMetaboliteNetwork(graph)
```

```{r echo=TRUE}
#fix attr of vertices in mgraph object
mgraph_V_attr <- unlist(V(mgraph)$attr)
mgraph <- set_vertex_attr(graph= mgraph, name= "miriam.kegg.compound", value= mgraph_V_attr)
mgraph <- delete_vertex_attr(graph= mgraph, name= "attr")
#View(igraph::as_data_frame(mgraph, "vertices"))
```

```{r echo=TRUE}
# fix attr of edges in mgraph object
mgraph_E_attr <- as.data.frame(do.call(rbind, E(mgraph)$attr))
for(c in colnames(mgraph_E_attr)){
  mgraph <- set_edge_attr(graph= mgraph, name = c, value = mgraph_E_attr[[c]])
}

mgraph <- delete_edge_attr(graph= mgraph, name= "attr")
mgraph <- delete_edge_attr(graph= mgraph, name= "reactant.stoichiometry")
mgraph <- delete_edge_attr(graph= mgraph, name= "product.stoichiometry")
mgraph <- delete_edge_attr(graph= mgraph, name= "kinetics")

View(igraph::as_data_frame(mgraph))
View(igraph::as_data_frame(mgraph, "vertices"))
```

As we can see, the network contains isolated structures. We can find out, how much isolated "subgraphs" exist:

```{r echo=FALSE}
vn.mgraph <- toVisNetworkData(mgraph)
visNetwork(vn.mgraph$nodes, vn.mgraph$edges) %>%  visEdges(arrows = 'to', smooth =T)
```

```{r echo=TRUE}
#components(mgraph)
split(names(V(mgraph)), components(mgraph)$membership)
```

The two big isolated subnetworks (1 & 2) are caused because there are two different IDs for Lactosylceramide (C01290 & G00092). We merge the two vertices C01290 & G00092 and finally remove the vertex G00092. 

```{r}
#V(mgraph)$name <- as.numeric(seq(1, length(V(mgraph))))
nodes <- 1:vcount(mgraph)
nodes[16] <- 36

mgraph <- contract.vertices(graph=mgraph, mapping=nodes, vertex.attr.comb=list("first"))
mgraph <- delete.vertices(mgraph, 16)

V(mgraph)[[16]]$name <- 36
V(mgraph)[[16]]$miriam.kegg.compound <- "00092"

View(igraph::as_data_frame(mgraph))
View(igraph::as_data_frame(mgraph, "vertices"))
```

```{r echo=FALSE}
vn.mgraph <- toVisNetworkData(mgraph)
visNetwork(vn.mgraph$nodes, vn.mgraph$edges) %>%  visEdges(arrows = 'to', smooth =T)
```

```{r echo=TRUE}
#components(mgraph)
split(names(V(mgraph)), components(mgraph)$membership)
```


The two big isolated subnetworks (1 & 2) are caused because there are two different IDs for Lactosylceramide (C01290 & G00092). In the following step, we connect these two subnetworks by 
* removing the node C01290 (LacCer) and 
* connecting C01190 (Glucosylceramide) with G00092 (LacCer) by adding two new edges (identical to the removed edges between C01290 and C01190)
* TODO: connecting C03405 (Lactosylceramide sulfate) with G00092 (LacCer) by adding two new edges (identical to the removed edges between C03405 and C01190)

**Number of verices:**
```{r echo=FALSE}
# number of verices
length(V(mgraph))
```

**Number of edges:**
```{r echo=FALSE}
# number of edges
length(E(mgraph))
```

Use edge attribute `miriam.ncbigene` to map the NCBI gene IDs to `SYMBOL` and `ENSEMBLE` with the package `org.Hs.eg.db`.

```{r results=FALSE}
# 
# edge_attr(graph = mgraph, name = "symbol") <- list(list())
# edge_attr(graph = mgraph, name = "ensembl") <- list(list())
for(e in seq(1:length(E(mgraph)))){
  # print(edge_attr(graph = mgraph, name = "miriam.ncbigene", index = e)[[1]])
  miriam.ncbigene <- edge_attr(graph = mgraph, name = "miriam.ncbigene", index = e)[[1]]
  if(!is.null(miriam.ncbigene)){
    miriam.symbol <- mapIds(org.Hs.eg.db, keys = miriam.ncbigene, keytype = "ENTREZID", column="SYMBOL")
    print(miriam.symbol)
    miriam.ensembl <- mapIds(org.Hs.eg.db, keys = miriam.ncbigene, keytype = "ENTREZID", column="ENSEMBL")
    print(miriam.ensembl)
  }
  edge_attr(graph = mgraph, name = "symbol", index = e) <- list(miriam.symbol)
  edge_attr(graph = mgraph, name = "ensembl", index = e) <- list(miriam.ensembl)
}
```

**Overview of the edges including their attributes:**

```{r echo=FALSE}
DT::datatable(as_data_frame(mgraph))
```

**Assign readable names to vertices**

```{r echo=TRUE}
labels <- sapply(names(V(mgraph)), function(x){
  l <- keggFind("glycan", x)
  if(rlang::is_empty(l)){
    l <- keggFind("compound", x)
  }
  unlist(strsplit(l, ";"))[1]
})

if(length(V(mgraph)) == length(labels)){
  mgraph <- set.vertex.attribute(mgraph, "label", value=labels)
  DT::datatable(as_data_frame(mgraph, "vertices")[,-c(1)])
}

# save for later use
saveRDS(mgraph, "./data/substrate_graph.Rds")
```

```{r echo=FALSE}
vn.mgraph <- toVisNetworkData(mgraph, idToLabel = FALSE)
visNetwork(vn.mgraph$nodes, vn.mgraph$edges) %>%  visEdges(arrows = 'to', smooth =T)
```

## Identify target features

```{r echo=TRUE}
target_features_df <- data.frame(
  Ensembl = unique(unlist(edge_attr(mgraph, "ensembl"))),
  Symbol = unique(unlist(edge_attr(mgraph, "symbol"))),
  NCBI_IDs = unique(unlist(edge_attr(mgraph, "miriam.ncbigene"))))

# save for later purpose
#saveRDS(target_features_df, "../data/metabolic_graphs/target_features.Rds")

DT::datatable(target_features_df)
```


