---
title: "GN & Neuroblastoma - pathanalysis"
author:
- name: Arsenij Ustjanzew (arsenij.ustjanzew@uni-mainz.de)<br><a href='https://www.unimedizin-mainz.de/imbei/'>IMBEI,
    University Medical Center Mainz</a><br>
date: "22.02.2023"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: hide
    code_download: true
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load necessary packages

```{r echo=TRUE, warning=FALSE, message=FALSE, results='hide'}
setwd("/home/rstudio/")
library("SummarizedExperiment")
library("DT")
library("DESeq2")
library("org.Hs.eg.db")
library("pheatmap")
library("dplyr")
#devtools::install_github("vqv/ggbiplot")
library("ggbiplot")
library("KEGGREST")
library("viridis")
library("igraph")
library("visNetwork")
library("EnhancedVolcano")
library("magrittr")

source("edge_probability_functions.R")
```

# Data set analysis: GN vs. Neuroblastoma


# Read data

```{r}
# load graph
mgraph <- readRDS("./data/substrate_graph.Rds")

# delete edges HEX-A/B 

mgraph %<>% delete_edges("00109|00108")
mgraph %<>% delete_edges("00110|00109")

vn.mgraph <- toVisNetworkData(mgraph, idToLabel = FALSE)
vn.mgraph$edges$id <- 1:nrow(vn.mgraph$edges)

# 118 edges (reaction R01281 is duplicated once)
# edge_df <- igraph::as_data_frame(mgraph)
```


```{r warning=FALSE, message=FALSE}
counts_dir <- "./data/GSE147635-counts"
count_files <- list.files(counts_dir)

# initialize an empty data.frame
file_path <- file.path(counts_dir, count_files[1])
df <- read.table(file_path, header = FALSE, sep = "\t")
names(df) <- c("rowname", count_files[1])

# loop through each file and bind it to the data.frame
for (file in count_files[2:length(count_files)]) {
  file_path <- file.path(counts_dir, file)
  print(file_path)
  temp_df <- read.table(file_path, header = FALSE, sep = "\t")
  names(temp_df) <- c("rowname", file)
  # rownames(temp_df) <- temp_df[,"rowname"]
  # print(head(temp_df))
  df <- cbind(df, temp_df[,file, drop=FALSE])
}
```

# Create count matrix, colData & rowData

```{r warning=FALSE, message=FALSE}
# column "rowname" as rownames
rownames(df) <- df$rowname
df <- df[,-1]

# fix colnames
colnames(df)  <- gsub("_count.tsv","",colnames(df))

# remove NB_CL samples
NB_CL <- c("SRR11434496", "SRR11434497", "SRR11434498")
df <- df[,-which(colnames(df)==NB_CL)]

#coldata
coldata_NB_GN <- data.frame(type=c(rep("GN", 6), rep("NB", 15)))
rownames(coldata_NB_GN) <- colnames(df)

#rowdata
ensemblIDs <- gsub("\\..*", "", rownames(df))
geneSymbol <- AnnotationDbi::mapIds(org.Hs.eg.db, keys = ensemblIDs, keytype = "ENSEMBL", column="SYMBOL")
entrezIDs <- AnnotationDbi::mapIds(org.Hs.eg.db, keys = ensemblIDs, keytype = "ENSEMBL", column="ENTREZID")

rowdata <- data.frame(ensemblIDs = ensemblIDs, geneSymbol = geneSymbol, entrezIDs = entrezIDs, row.names = rownames(df))

all(colnames(df) == rownames(coldata_NB_GN))
```


```{r}
# create DDS object
NB_GN_dds <- DESeq2::DESeqDataSetFromMatrix(df, coldata_NB_GN, design=~type)
rowData(NB_GN_dds) <- rowdata


# gene symbol as row name
NB_GN_dds <- NB_GN_dds[which(!is.na(rowData(NB_GN_dds)$geneSymbol)),]
NB_GN_dds <- NB_GN_dds[-duplicated(rowData(NB_GN_dds)$geneSymbol),]
rownames(NB_GN_dds) <- rowData(NB_GN_dds)$geneSymbol
```

Filtering genes with less then 10 counts in sum for all samples

```{r}
keep <- rowSums(counts(NB_GN_dds)) >= 10
NB_GN_dds <- NB_GN_dds[keep,]
```

# Normalization & log10 +1

```{r}
NB_GN_dds <- estimateSizeFactors(NB_GN_dds)

norm_counts <- DESeq2::counts(NB_GN_dds, normalized=TRUE)
norm_counts <- log10(norm_counts + 1)

saveRDS(NB_GN_dds, "NB_GN_dds_GSE147635.Rds")
```

# Distance matrix

```{r, fig.height= 18, , fig.width = 18}
vsd <- vst(NB_GN_dds, blind=FALSE)
sampleDists <- dist(t(assay(vsd)))

library("RColorBrewer")
sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- as.character(vsd$type)
colnames(sampleDistMatrix) <- NULL
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix)
```

# Assign gene expression to graph and compute reaction activity score (RAS)

```{r}
# create reaction activity matrix; weights are the sum of genes involved in the reaction
ras <- compute_reaction_activity(mgraph, norm_counts)

# create weighted graph per sample; weights are the sum of genes involved in the reaction
graph_list <- compute_reaction_activity(mgraph, norm_counts, output_graph=TRUE)
```

# Compute transition probability matrix

```{r}
# transition matrix (equal to random walker with max-steps 1)
transition_probability <- compute_transition_probablity(graph_list, "edge_sum", "miriam.kegg.reaction", target_node=NULL)

# transition matrix, where the product of the transition probabilities of all simple paths from the target_node are calculated  
transition_probability_paths <- compute_transition_probablity(graph_list, "edge_sum", "miriam.kegg.reaction", target_node="C01290")

# transition matrix, where if the transition probability = 1 is replaced by the recursively previous max. transition probability 
transition_probability_rec <- compute_transition_probablity(graph_list, "edge_sum", "miriam.kegg.reaction", target_node=NULL, pass_through=TRUE)
```

# Adjust RAS by transition probabilities

```{r}
r <- intersect(rownames(ras), rownames(transition_probability))
ras_prob <- ras[r,] * transition_probability[r,]
ras_prob_path <- ras[r,] * transition_probability_paths[r,]
ras_prob_rec <- ras[r,] * transition_probability_rec[r,]

ras_prob_up <- ras[r,] + (ras[r,] * transition_probability[r,])
ras_prob_up_path <- ras[r,] + (ras[r,] * transition_probability_paths[r,])
ras_prob_up_rec <- ras[r,] + (ras[r,] * transition_probability_rec[r,])
```

# RAS

```{r warning=FALSE, message=FALSE}
p.vals <- sapply(1:nrow(ras), function(x){
  ks.test(ras[x, which(colData(NB_GN_dds)$type=="NB")],ras[x, which(colData(NB_GN_dds)$type=="GN")])$p.value
})
pval_df <- data.frame(reaction=rownames(ras), pval=round(p.vals, 5), padj=p.adjust(p.vals, method="BH"))
log2fc <- sapply(1:nrow(ras), function(x){
  log2(mean(ras[x, which(colData(NB_GN_dds)$type=="NB")])) - log2(mean(ras[x, which(colData(NB_GN_dds)$type=="GN")]))
})
pval_df$log2fc <- log2fc
rownames(pval_df) <- gsub("| .*", "", pval_df$reaction)

pval_df$NB <- rowMeans(ras[, which(colData(NB_GN_dds)$type=="NB")])
pval_df$GN <- rowMeans(ras[, which(colData(NB_GN_dds)$type=="GN")])

# lower_ci_NB <- sapply(1:nrow(ras), function(x){
#   t.test(ras[x, which(colData(NB_GN_dds)$type=="NB")], conf.level = 0.95)$conf.int[1]
# })
# upper_ci_NB <- sapply(1:nrow(ras), function(x){
#   t.test(ras[x, which(colData(NB_GN_dds)$type=="NB")], conf.level = 0.95)$conf.int[2]
# })
# lower_ci_GN <- sapply(1:nrow(ras), function(x){
#   t.test(ras[x, which(colData(NB_GN_dds)$type=="GN")], conf.level = 0.95)$conf.int[1]
# })
# upper_ci_GN <- sapply(1:nrow(ras), function(x){
#   t.test(ras[x, which(colData(NB_GN_dds)$type=="GN")], conf.level = 0.95)$conf.int[2]
# })

pval_df$sd_NB <- sapply(1:nrow(ras), function(x){
  sd(ras[x, which(colData(NB_GN_dds)$type=="NB")])
})
pval_df$sd_GN <- sapply(1:nrow(ras), function(x){
  sd(ras[x, which(colData(NB_GN_dds)$type=="GN")])
})

writexl::write_xlsx(pval_df, "./data/NB_GN_FC_images/NB_GN_ras_tbl_FCstatistic.xlsx")
```

```{r fig.width = 10, fig.height=10}
vn.mgraph <- visNetwork::toVisNetworkData(mgraph, idToLabel = FALSE)
vn.mgraph$edges$id <- 1:nrow(vn.mgraph$edges)


vn.mgraph$edges$miriam.kegg.reaction <- unlist(vn.mgraph$edges$miriam.kegg.reaction)
vn.mgraph$edges <- merge(as.data.frame(vn.mgraph$edges), pval_df, by.x = which(colnames(vn.mgraph$edges) =="miriam.kegg.reaction"), by.y = 0, all.x = TRUE)

vn.mgraph$edges$color <- ifelse(vn.mgraph$edges$log2fc < 0, "blue", "red")
vn.mgraph$edges[which(vn.mgraph$edges$padj > 0.05),"color"] <- "grey"
vn.mgraph$edges$value <- abs(vn.mgraph$edges$log2fc)
vn.mgraph$edges$title <- round(vn.mgraph$edges$log2fc,3)

# visNetwork(vn.mgraph$nodes, vn.mgraph$edges) %>%  visEdges(arrows = 'to', smooth =T)
# 
# # compounds of gangliosphingolipid
# comps <- names(keggGet("hsa00604")[[1]]$COMPOUND)
# comps <- gsub("G", "", comps)
# comps[1] <- "C01290"
# 
# vn.mgraph$nodes[which(vn.mgraph$nodes$miriam.kegg.compound == "00127"),"label"] <- "GD1alpha"
# 
# vn.mgraph$nodes <- vn.mgraph$nodes[which(vn.mgraph$nodes$id %in% comps),]
# vn.mgraph$nodes$level <- c(
#   "Lactosylceramide" = 1,
#   "GM1" = 3,
#   "GM2" = 2,
#   "GD3" = 1,
#   "GT2" = 2,
#   "GT3" = 1,
#   "GD1b" = 3,
#   "GT1b" = 4,
#   "GD1a" = 4,
#   "GM1b" = 4,
#   "GM3" = 1,
#   "GD2" = 2,
#   "GA2" = 2,
#   "GA1" = 3,
#   "GT1c" = 3,
#   "GQ1b" = 5,
#   "GQ1balpha" = 5,
#   "GT1aalpha" = 5,
#   "GT1a" = 5,
#   "GD1c" = 5,
#   "GD1a1" = 4)
# # 
# visNetwork(vn.mgraph$nodes, vn.mgraph$edges) %>%
#   visEdges(arrows = 'to', smooth =FALSE, arrowStrikethrough = FALSE, font = list(size=14, align="bottom")) %>%  
#   visNodes(shape = 'box') %>% 
#   visHierarchicalLayout(levelSeparation = NULL, nodeSpacing = 1, edgeMinimization=TRUE,treeSpacing=1) %>%
#   visPhysics(enabled = FALSE)

```

```{r}
create_graph <- function(mgraph){
  # compounds of gangliosphingolipid
  comps <- names(keggGet("hsa00604")[[1]]$COMPOUND)
  comps <- gsub("G", "", comps)
  comps[1] <- "C01290"
  
  mgraph_sub <- subgraph(mgraph, which(V(mgraph)$name %in% comps))
  
  V(mgraph_sub)$label[which(V(mgraph_sub)$miriam.kegg.compound == "00127")] <- "GD1alpha"
  
  mylay <- layout.auto(mgraph_sub)
  rownames(mylay) <- V(mgraph_sub)$name
  
  mylay["C01290",] <-c(-6,0)
  mylay["00108",] <-c(4,0)
  mylay["00113",] <-c(12,0)
  mylay["00118",] <-c(20,0)
  mylay["00123",] <-c(-4,-4)
  mylay["00109",] <-c(4,-4)
  mylay["00114",] <-c(12,-4)
  mylay["00119",] <-c(20,-4)
  mylay["00124",] <-c(-4,-8)
  mylay["00110",] <-c(4,-8)
  mylay["00115",] <-c(12,-8)
  mylay["00120",] <-c(20,-8)
  mylay["00125",] <-c(-4,-12)
  mylay["00111",] <-c(4,-12)
  mylay["00116",] <-c(12,-12)
  mylay["00127",] <-c(-6.2,-18)
  mylay["00126",] <-c(-1.2,-16)
  mylay["00128",] <-c(1.8,-18)
  mylay["00112",] <-c(6.2,-16)
  mylay["00129",] <-c(9.8,-18)
  mylay["00117",] <-c(14.2,-16)
  
  V(mgraph_sub)$shape <- "rectangle"
  V(mgraph_sub)$size <- 100
  V(mgraph_sub)$size <- 35
  V(mgraph_sub)["C01290"]$size <- 75
  V(mgraph_sub)["00128"]$size <- 50
  V(mgraph_sub)["00129"]$size <- 50
  V(mgraph_sub)["00127"]$size <- 50
  V(mgraph_sub)$label.cex <- 1
  V(mgraph_sub)$color <- "lightblue"
  r.id <- unlist(get.edge.attribute(mgraph_sub,'miriam.kegg.reaction'))
  mgraph_sub <- set_edge_attr(mgraph_sub, "log2fc", index = E(mgraph_sub), pval_df[r.id,"log2fc"])
  mgraph_sub <- set_edge_attr(mgraph_sub, "color", index = E(mgraph_sub), "NA")
  mgraph_sub <- set_edge_attr(mgraph_sub, "padj", index = E(mgraph_sub), pval_df[r.id,"padj"])
  E(mgraph_sub)$width <- abs(as.numeric(E(mgraph_sub)$log2fc)) * 40
  
  E(mgraph_sub)$color <- ifelse(E(mgraph_sub)$log2fc < 0, "blue", "red")
  # E(mgraph_sub)$color <- ifelse(E(mgraph_sub)$log2fc < 0, "blue", "red")
  E(mgraph_sub)$color[which(E(mgraph_sub)$padj > 0.05)] <- "grey"
  return(mgraph_sub)
}
```


```{r}
mgraph_sub <- create_graph(mgraph)
# sapply(E(mgraph_sub)$log2fc, is.greaterFoldChange)
png("./data/NB_GN_FC_images/NB_GN_graph_ras.png", width = 2800, height = 2800, units = "px", pointsize = 20, res="300")
par(omi=c(0,0,0,0), mgp=c(0,0,0),mar=c(0,0,0,0))
plot(mgraph_sub, layout=mylay, vertex.label.color= "black", vertex.label.family= "Arial")
dev.off()
```

```{r}
e_df <- as.data.frame(vn.mgraph$edges)

# ggplot(e_df, aes(x=1, y=miriam.kegg.reaction)) +
#   geom_point(aes(size=-log10(padj), color=log2fc)) +
#   colorspace::scale_color_continuous_diverging(palette = "Blue-Red 3") +
#   theme_classic()
# 
# 
# 
# 
# ggplot(e_df, aes(x=as.character(to), y=as.character(from), fill=log2fc)) +
#   geom_point(aes(size=-log10(padj), shape = 21, colour = "black")) +
#   colorspace::scale_color_continuous_diverging(palette = "Blue-Red 3") +
#   theme_classic() + 
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
#   xlab("from node") + ylab("to node")



# ggplot(e_df, aes(x=1, y=miriam.kegg.reaction)) +
#   geom_point(aes(size=-log10(padj), color=log2fc)) +
#   colorspace::scale_color_continuous_diverging(palette = "Blue-Red 3") +
#   theme_classic()
# 
# 
# 
# 
# ggplot(e_df, aes(x=as.character(to), y=as.character(from), fill=log2fc)) +
#   geom_point(aes(size=-log10(padj), shape = 21, colour = "black")) +
#   colorspace::scale_color_continuous_diverging(palette = "Blue-Red 3") +
#   theme_classic() + 
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
#   xlab("from node") + ylab("to node")

e_df_lg <- e_df %>%
  tidyr::pivot_longer(colnames(dplyr::select(e_df, c("NB", "GN", "log2fc"))), names_to = "group", values_to = "mean")
e_df_lg$group <- as.factor(e_df_lg$group)
e_df_lg[which(e_df_lg$group %in% c("NB", "GN")),"padj"] <- NA
e_df_lg[which(e_df_lg$group %in% c("log2fc")),"sd_NB"] <- NA
e_df_lg[which(e_df_lg$group %in% c("log2fc")),"sd_GN"] <- NA
e_df_lg[which(e_df_lg$group == "GN"),"sd_NB"] <- NA
e_df_lg[which(e_df_lg$group == "NB"),"sd_GN"] <- NA

e_df_lg <- e_df_lg %>% mutate(sd = coalesce(sd_NB,sd_GN))  

# ggplot(e_df_lg, aes(x=mean, y=miriam.kegg.reaction, group=group)) + 
#   geom_point(aes(color=-log10(padj))) + 
#   colorspace::scale_color_continuous_sequential(palette = "Reds 3") +
#   facet_wrap(~group, ncol = 3) + 
#   theme_classic() + 
#   theme(panel.grid.major.y = element_line(color = "gray",
#                                           size = 0.5,
#                                           linetype = 1)) + xlab("")


comps <- names(keggGet("hsa00604")[[1]]$COMPOUND)
comps <- gsub("G", "", comps)
comps[1] <- "C01290"
comps <- c(comps, "C01190")

e_df_lg <- e_df_lg[which(e_df_lg$from %in% comps | e_df_lg$to %in% comps),]
e_df_lg$geneSymbol <- sapply(e_df_lg$symbol, function(x) paste0(unique(x), collapse = ' '))
from_label <- c()
for(i in e_df_lg$from){
  from_label <- c(from_label, vn.mgraph$nodes[which(vn.mgraph$nodes$id == i), "label"])
}
e_df_lg$from_label <- from_label
to_label <- c()
for(i in e_df_lg$to){
  to_label <- c(to_label, vn.mgraph$nodes[which(vn.mgraph$nodes$id == i), "label"])
}
e_df_lg$to_label <- to_label

# ggplot(e_df_lg, aes(x=mean, y=paste0(miriam.kegg.reaction, " (", geneSymbol, ")"), group=group)) + 
  
p1 <- ggplot(e_df_lg, aes(x=mean, y=paste0(miriam.kegg.reaction, " (from: ", from_label, " - to: ",to_label, ")"), group=group)) + 
  geom_vline(xintercept = 0, colour="lightgray", linetype="solid") +
  geom_errorbar(aes(xmin=mean - sd,xmax=mean + sd,width=0.2)) +
  geom_point(aes(color=-log10(padj))) +
  colorspace::scale_color_continuous_sequential(palette = "Reds 3") +
  facet_wrap(~group, ncol = 3, scales="free_x") + 
  theme_classic() +
  ylab("") + xlab("") +
  theme(panel.grid.major.y = element_line(color = "lightgray",
                                          size = 0.5,
                                          linetype = 1)) + xlab("")
ggsave(
  "NB_GN_FC_ras.png",
  plot = p1,
  device = "png",
  path = "./data/NB_GN_FC_images/",
  scale = 1,
  width = 2800,
  height = 1500,
  units = "px",
  dpi = 300,
  limitsize = FALSE,
  bg = NULL,
)
```

# RAS prob

```{r warning=FALSE, message=FALSE}
p.vals <- sapply(1:nrow(ras_prob), function(x){
  # print(ras_prob[x, which(colData(NB_GN_dds)$type=="NB")])
  ks.test(ras_prob[x, which(colData(NB_GN_dds)$type=="NB")],ras_prob[x, which(colData(NB_GN_dds)$type=="GN")])$p.value
})
pval_df <- data.frame(reaction=rownames(ras_prob), pval=round(p.vals, 5), padj=p.adjust(p.vals, method="BH"))
log2fc <- sapply(1:nrow(ras_prob), function(x){
  log2(mean(ras_prob[x, which(colData(NB_GN_dds)$type=="NB")])) - log2(mean(ras_prob[x, which(colData(NB_GN_dds)$type=="GN")]))
})
pval_df$log2fc <- log2fc
rownames(pval_df) <- gsub("| .*", "", pval_df$reaction)

pval_df$NB <- rowMeans(ras_prob[, which(colData(NB_GN_dds)$type=="NB")])
pval_df$GN <- rowMeans(ras_prob[, which(colData(NB_GN_dds)$type=="GN")])

# lower_ci_NB <- sapply(1:nrow(ras_prob), function(x){
#   t.test(ras_prob[x, which(colData(NB_GN_dds)$type=="NB")], conf.level = 0.95)$conf.int[1]
# })
# upper_ci_NB <- sapply(1:nrow(ras_prob), function(x){
#   t.test(ras_prob[x, which(colData(NB_GN_dds)$type=="NB")], conf.level = 0.95)$conf.int[2]
# })
# lower_ci_GN <- sapply(1:nrow(ras_prob), function(x){
#   t.test(ras_prob[x, which(colData(NB_GN_dds)$type=="GN")], conf.level = 0.95)$conf.int[1]
# })
# upper_ci_GN <- sapply(1:nrow(ras_prob), function(x){
#   t.test(ras_prob[x, which(colData(NB_GN_dds)$type=="GN")], conf.level = 0.95)$conf.int[2]
# })

pval_df$sd_NB <- sapply(1:nrow(ras_prob), function(x){
  sd(ras_prob[x, which(colData(NB_GN_dds)$type=="NB")])
})
pval_df$sd_GN <- sapply(1:nrow(ras_prob), function(x){
  sd(ras_prob[x, which(colData(NB_GN_dds)$type=="GN")])
})

writexl::write_xlsx(pval_df, "./data/NB_GN_FC_images/NB_GN_ras_prob_tbl_FCstatistic.xlsx")
```

```{r fig.width = 10, fig.height=10}
vn.mgraph <- visNetwork::toVisNetworkData(mgraph, idToLabel = FALSE)
vn.mgraph$edges$id <- 1:nrow(vn.mgraph$edges)


vn.mgraph$edges$miriam.kegg.reaction <- unlist(vn.mgraph$edges$miriam.kegg.reaction)
vn.mgraph$edges <- merge(as.data.frame(vn.mgraph$edges), pval_df, by.x = which(colnames(vn.mgraph$edges) =="miriam.kegg.reaction"), by.y = 0, all.x = TRUE)

vn.mgraph$edges$color <- ifelse(vn.mgraph$edges$log2fc < 0, "blue", "red")
vn.mgraph$edges[which(vn.mgraph$edges$padj > 0.05),"color"] <- "grey"
vn.mgraph$edges$value <- abs(vn.mgraph$edges$log2fc)
vn.mgraph$edges$title <- round(vn.mgraph$edges$log2fc,3)

# visNetwork(vn.mgraph$nodes, vn.mgraph$edges) %>%  visEdges(arrows = 'to', smooth =T)
# 
# # compounds of gangliosphingolipid
# comps <- names(keggGet("hsa00604")[[1]]$COMPOUND)
# comps <- gsub("G", "", comps)
# comps[1] <- "C01290"
# 
# vn.mgraph$nodes[which(vn.mgraph$nodes$miriam.kegg.compound == "00127"),"label"] <- "GD1alpha"
# 
# vn.mgraph$nodes <- vn.mgraph$nodes[which(vn.mgraph$nodes$id %in% comps),]
# vn.mgraph$nodes$level <- c(
#   "Lactosylceramide" = 1,
#   "GM1" = 3,
#   "GM2" = 2,
#   "GD3" = 1,
#   "GT2" = 2,
#   "GT3" = 1,
#   "GD1b" = 3,
#   "GT1b" = 4,
#   "GD1a" = 4,
#   "GM1b" = 4,
#   "GM3" = 1,
#   "GD2" = 2,
#   "GA2" = 2,
#   "GA1" = 3,
#   "GT1c" = 3,
#   "GQ1b" = 5,
#   "GQ1balpha" = 5,
#   "GT1aalpha" = 5,
#   "GT1a" = 5,
#   "GD1c" = 5,
#   "GD1a1" = 4)
# # 
# visNetwork(vn.mgraph$nodes, vn.mgraph$edges) %>%
#   visEdges(arrows = 'to', smooth =FALSE, arrowStrikethrough = FALSE, font = list(size=14, align="bottom")) %>%  
#   visNodes(shape = 'box') %>% 
#   visHierarchicalLayout(levelSeparation = NULL, nodeSpacing = 1, edgeMinimization=TRUE,treeSpacing=1) %>%
#   visPhysics(enabled = FALSE)

```

```{r}
mgraph_sub <- create_graph(mgraph)

png("./data/NB_GN_FC_images/NB_GN_graph_ras_prob.png", width = 2800, height = 2800, units = "px", pointsize = 20, res="300")
par(omi=c(0,0,0,0), mgp=c(0,0,0),mar=c(0,0,0,0))
plot(mgraph_sub, layout=mylay, vertex.label.color= "black", vertex.label.family= "Arial")
dev.off()
```

```{r}
e_df <- as.data.frame(vn.mgraph$edges)

# ggplot(e_df, aes(x=1, y=miriam.kegg.reaction)) +
#   geom_point(aes(size=-log10(padj), color=log2fc)) +
#   colorspace::scale_color_continuous_diverging(palette = "Blue-Red 3") +
#   theme_classic()
# 
# 
# 
# 
# ggplot(e_df, aes(x=as.character(to), y=as.character(from), fill=log2fc)) +
#   geom_point(aes(size=-log10(padj), shape = 21, colour = "black")) +
#   colorspace::scale_color_continuous_diverging(palette = "Blue-Red 3") +
#   theme_classic() + 
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
#   xlab("from node") + ylab("to node")

e_df_lg <- e_df %>%
  tidyr::pivot_longer(colnames(dplyr::select(e_df, c("NB", "GN", "log2fc"))), names_to = "group", values_to = "mean")
e_df_lg$group <- as.factor(e_df_lg$group)
e_df_lg[which(e_df_lg$group %in% c("NB", "GN")),"padj"] <- NA
e_df_lg[which(e_df_lg$group %in% c("log2fc")),"sd_NB"] <- NA
e_df_lg[which(e_df_lg$group %in% c("log2fc")),"sd_GN"] <- NA
e_df_lg[which(e_df_lg$group == "GN"),"sd_NB"] <- NA
e_df_lg[which(e_df_lg$group == "NB"),"sd_GN"] <- NA

e_df_lg <- e_df_lg %>% mutate(sd = coalesce(sd_NB,sd_GN)) 

# ggplot(e_df_lg, aes(x=mean, y=miriam.kegg.reaction, group=group)) + 
#   geom_point(aes(color=-log10(padj))) + 
#   colorspace::scale_color_continuous_sequential(palette = "Reds 3") +
#   facet_wrap(~group, ncol = 3) + 
#   theme_classic() + 
#   theme(panel.grid.major.y = element_line(color = "gray",
#                                           size = 0.5,
#                                           linetype = 1)) + xlab("")


comps <- names(keggGet("hsa00604")[[1]]$COMPOUND)
comps <- gsub("G", "", comps)
comps[1] <- "C01290"
comps <- c(comps, "C01190")

e_df_lg <- e_df_lg[which(e_df_lg$from %in% comps | e_df_lg$to %in% comps),]
e_df_lg$geneSymbol <- sapply(e_df_lg$symbol, function(x) paste0(unique(x), collapse = ' '))
from_label <- c()
for(i in e_df_lg$from){
  from_label <- c(from_label, vn.mgraph$nodes[which(vn.mgraph$nodes$id == i), "label"])
}
e_df_lg$from_label <- from_label
to_label <- c()
for(i in e_df_lg$to){
  to_label <- c(to_label, vn.mgraph$nodes[which(vn.mgraph$nodes$id == i), "label"])
}
e_df_lg$to_label <- to_label

# ggplot(e_df_lg, aes(x=mean, y=paste0(miriam.kegg.reaction, " (", geneSymbol, ")"), group=group)) + 
  
p2 <- ggplot(e_df_lg, aes(x=mean, y=paste0(miriam.kegg.reaction, " (from: ", from_label, " - to: ",to_label, ")"), group=group)) + 
  geom_vline(xintercept = 0, colour="lightgray", linetype="solid") +
  geom_errorbar(aes(xmin=mean - sd,xmax=mean + sd,width=0.2)) +
  geom_point(aes(color=-log10(padj))) +
  colorspace::scale_color_continuous_sequential(palette = "Reds 3") +
  facet_wrap(~group, ncol = 3, scales="free_x") + 
  theme_classic() +
  ylab("") + xlab("") +
  theme(panel.grid.major.y = element_line(color = "lightgray",
                                          size = 0.5,
                                          linetype = 1)) + xlab("")

ggsave(
  "NB_GN_FC_ras_prob.png",
  plot = p2,
  device = "png",
  path = "./data/NB_GN_FC_images/",
  scale = 1,
  width = 2800,
  height = 1500,
  units = "px",
  dpi = 300,
  limitsize = FALSE,
  bg = NULL,
)
```

# RAS prob rec

```{r warning=FALSE, message=FALSE}
p.vals <- sapply(1:nrow(ras_prob_rec), function(x){
  # print(ras_prob_rec[x, which(colData(NB_GN_dds)$type=="NB")])
  ks.test(ras_prob_rec[x, which(colData(NB_GN_dds)$type=="NB")],ras_prob_rec[x, which(colData(NB_GN_dds)$type=="GN")])$p.value
})
pval_df <- data.frame(reaction=rownames(ras_prob_rec), pval=round(p.vals, 5), padj=p.adjust(p.vals, method="BH"))
log2fc <- sapply(1:nrow(ras_prob_rec), function(x){
  log2(mean(ras_prob_rec[x, which(colData(NB_GN_dds)$type=="NB")])) - log2(mean(ras_prob_rec[x, which(colData(NB_GN_dds)$type=="GN")]))
})
pval_df$log2fc <- log2fc
rownames(pval_df) <- gsub("| .*", "", pval_df$reaction)

pval_df$NB <- rowMeans(ras_prob_rec[, which(colData(NB_GN_dds)$type=="NB")])
pval_df$GN <- rowMeans(ras_prob_rec[, which(colData(NB_GN_dds)$type=="GN")])
# lower_ci_NB <- sapply(1:nrow(ras_prob_rec), function(x){
#   t.test(ras_prob_rec[x, which(colData(NB_GN_dds)$type=="NB")], conf.level = 0.95)$conf.int[1]
# })
# upper_ci_NB <- sapply(1:nrow(ras_prob_rec), function(x){
#   t.test(ras_prob_rec[x, which(colData(NB_GN_dds)$type=="NB")], conf.level = 0.95)$conf.int[2]
# })
# lower_ci_GN <- sapply(1:nrow(ras_prob_rec), function(x){
#   t.test(ras_prob_rec[x, which(colData(NB_GN_dds)$type=="GN")], conf.level = 0.95)$conf.int[1]
# })
# upper_ci_GN <- sapply(1:nrow(ras_prob_rec), function(x){
#   t.test(ras_prob_rec[x, which(colData(NB_GN_dds)$type=="GN")], conf.level = 0.95)$conf.int[2]
# })

pval_df$sd_NB <- sapply(1:nrow(ras_prob_rec), function(x){
  sd(ras_prob_rec[x, which(colData(NB_GN_dds)$type=="NB")])
})
pval_df$sd_GN <- sapply(1:nrow(ras_prob_rec), function(x){
  sd(ras_prob_rec[x, which(colData(NB_GN_dds)$type=="GN")])
})

writexl::write_xlsx(pval_df, "./data/NB_GN_FC_images/NB_GN_ras_prob_rec_tbl_FCstatistic.xlsx")
```

```{r fig.width = 10, fig.height=10}
vn.mgraph <- visNetwork::toVisNetworkData(mgraph, idToLabel = FALSE)
vn.mgraph$edges$id <- 1:nrow(vn.mgraph$edges)


vn.mgraph$edges$miriam.kegg.reaction <- unlist(vn.mgraph$edges$miriam.kegg.reaction)
vn.mgraph$edges <- merge(as.data.frame(vn.mgraph$edges), pval_df, by.x = which(colnames(vn.mgraph$edges) =="miriam.kegg.reaction"), by.y = 0, all.x = TRUE)

vn.mgraph$edges$color <- ifelse(vn.mgraph$edges$log2fc < 0, "blue", "red")
vn.mgraph$edges[which(vn.mgraph$edges$padj > 0.05),"color"] <- "grey"
vn.mgraph$edges$value <- abs(vn.mgraph$edges$log2fc)
vn.mgraph$edges$title <- round(vn.mgraph$edges$log2fc,3)

# visNetwork(vn.mgraph$nodes, vn.mgraph$edges) %>%  visEdges(arrows = 'to', smooth =T)
# 
# # compounds of gangliosphingolipid
# comps <- names(keggGet("hsa00604")[[1]]$COMPOUND)
# comps <- gsub("G", "", comps)
# comps[1] <- "C01290"
# 
# vn.mgraph$nodes[which(vn.mgraph$nodes$miriam.kegg.compound == "00127"),"label"] <- "GD1alpha"
# 
# vn.mgraph$nodes <- vn.mgraph$nodes[which(vn.mgraph$nodes$id %in% comps),]
# vn.mgraph$nodes$level <- c(
#   "Lactosylceramide" = 1,
#   "GM1" = 3,
#   "GM2" = 2,
#   "GD3" = 1,
#   "GT2" = 2,
#   "GT3" = 1,
#   "GD1b" = 3,
#   "GT1b" = 4,
#   "GD1a" = 4,
#   "GM1b" = 4,
#   "GM3" = 1,
#   "GD2" = 2,
#   "GA2" = 2,
#   "GA1" = 3,
#   "GT1c" = 3,
#   "GQ1b" = 5,
#   "GQ1balpha" = 5,
#   "GT1aalpha" = 5,
#   "GT1a" = 5,
#   "GD1c" = 5,
#   "GD1a1" = 4)
# # 
# visNetwork(vn.mgraph$nodes, vn.mgraph$edges) %>%
#   visEdges(arrows = 'to', smooth =FALSE, arrowStrikethrough = FALSE, font = list(size=14, align="bottom")) %>%  
#   visNodes(shape = 'box') %>% 
#   visHierarchicalLayout(levelSeparation = NULL, nodeSpacing = 1, edgeMinimization=TRUE,treeSpacing=1) %>%
#   visPhysics(enabled = FALSE)

```

```{r}
mgraph_sub <- create_graph(mgraph)

png("./data/NB_GN_FC_images/NB_GN_graph_ras_prob_rec.png", width = 2800, height = 2800, units = "px", pointsize = 20, res="300")
par(omi=c(0,0,0,0), mgp=c(0,0,0),mar=c(0,0,0,0))
plot(mgraph_sub, layout=mylay, vertex.label.color= "black", vertex.label.family= "Arial")
dev.off()
```

```{r}
e_df <- as.data.frame(vn.mgraph$edges)

# ggplot(e_df, aes(x=1, y=miriam.kegg.reaction)) +
#   geom_point(aes(size=-log10(padj), color=log2fc)) +
#   colorspace::scale_color_continuous_diverging(palette = "Blue-Red 3") +
#   theme_classic()
# 
# 
# 
# 
# ggplot(e_df, aes(x=as.character(to), y=as.character(from), fill=log2fc)) +
#   geom_point(aes(size=-log10(padj), shape = 21, colour = "black")) +
#   colorspace::scale_color_continuous_diverging(palette = "Blue-Red 3") +
#   theme_classic() + 
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
#   xlab("from node") + ylab("to node")

e_df_lg <- e_df %>%
  tidyr::pivot_longer(colnames(dplyr::select(e_df, c("NB", "GN", "log2fc"))), names_to = "group", values_to = "mean")
e_df_lg$group <- as.factor(e_df_lg$group)
e_df_lg[which(e_df_lg$group %in% c("NB", "GN")),"padj"] <- NA
e_df_lg[which(e_df_lg$group %in% c("log2fc")),"sd_NB"] <- NA
e_df_lg[which(e_df_lg$group %in% c("log2fc")),"sd_GN"] <- NA
e_df_lg[which(e_df_lg$group == "GN"),"sd_NB"] <- NA
e_df_lg[which(e_df_lg$group == "NB"),"sd_GN"] <- NA

e_df_lg <- e_df_lg %>% mutate(sd = coalesce(sd_NB,sd_GN)) 

# ggplot(e_df_lg, aes(x=mean, y=miriam.kegg.reaction, group=group)) + 
#   geom_point(aes(color=-log10(padj))) + 
#   colorspace::scale_color_continuous_sequential(palette = "Reds 3") +
#   facet_wrap(~group, ncol = 3) + 
#   theme_classic() + 
#   theme(panel.grid.major.y = element_line(color = "gray",
#                                           size = 0.5,
#                                           linetype = 1)) + xlab("")


comps <- names(keggGet("hsa00604")[[1]]$COMPOUND)
comps <- gsub("G", "", comps)
comps[1] <- "C01290"
comps <- c(comps, "C01190")

e_df_lg <- e_df_lg[which(e_df_lg$from %in% comps | e_df_lg$to %in% comps),]
e_df_lg$geneSymbol <- sapply(e_df_lg$symbol, function(x) paste0(unique(x), collapse = ' '))
from_label <- c()
for(i in e_df_lg$from){
  from_label <- c(from_label, vn.mgraph$nodes[which(vn.mgraph$nodes$id == i), "label"])
}
e_df_lg$from_label <- from_label
to_label <- c()
for(i in e_df_lg$to){
  to_label <- c(to_label, vn.mgraph$nodes[which(vn.mgraph$nodes$id == i), "label"])
}
e_df_lg$to_label <- to_label

# ggplot(e_df_lg, aes(x=mean, y=paste0(miriam.kegg.reaction, " (", geneSymbol, ")"), group=group)) + 
  
p3 <- ggplot(e_df_lg, aes(x=mean, y=paste0(miriam.kegg.reaction, " (from: ", from_label, " - to: ",to_label, ")"), group=group)) + 
  geom_vline(xintercept = 0, colour="lightgray", linetype="solid") +
  geom_errorbar(aes(xmin=mean - sd,xmax=mean + sd,width=0.2)) +
  geom_point(aes(color=-log10(padj))) +
  colorspace::scale_color_continuous_sequential(palette = "Reds 3") +
  facet_wrap(~group, ncol = 3, scales="free_x") + 
  theme_classic() +
  ylab("") + xlab("") +
  theme(panel.grid.major.y = element_line(color = "lightgray",
                                          size = 0.5,
                                          linetype = 1)) + xlab("")

ggsave(
  "NB_GN_FC_ras_prob_rec.png",
  plot = p3,
  device = "png",
  path = "./data/NB_GN_FC_images/",
  scale = 1,
  width = 2800,
  height = 1500,
  units = "px",
  dpi = 300,
  limitsize = FALSE,
  bg = NULL,
)
```

# RAS prob path

```{r warning=FALSE, message=FALSE}

# ras_prob_path2 <- ras_prob_path[!apply(ras_prob_path, 1, function(x) all(x==0)), ]
ras_prob_path2 <- as.matrix(ras_prob_path)

p.vals <- sapply(1:nrow(ras_prob_path2), function(x){
  # print(ras_prob_path2[x, which(colData(NB_GN_dds)$type=="NB")])
  ks.test(ras_prob_path2[x, which(colData(NB_GN_dds)$type=="NB")],ras_prob_path2[x, which(colData(NB_GN_dds)$type=="GN")])$p.value
})
pval_df <- data.frame(reaction=rownames(ras_prob_path2), pval=round(p.vals, 5), padj=p.adjust(p.vals, method="BH"))
log2fc <- sapply(1:nrow(ras_prob_path2), function(x){
  log2(mean(ras_prob_path2[x, which(colData(NB_GN_dds)$type=="NB")])) - log2(mean(ras_prob_path2[x, which(colData(NB_GN_dds)$type=="GN")]))
})
pval_df$log2fc <- log2fc
rownames(pval_df) <- gsub("| .*", "", pval_df$reaction)

pval_df[which(is.nan(pval_df$log2fc)),"log2fc"] <- 0

pval_df$NB <- rowMeans(ras_prob_path2[, which(colData(NB_GN_dds)$type=="NB")])
pval_df$GN <- rowMeans(ras_prob_path2[, which(colData(NB_GN_dds)$type=="GN")])

# lower_ci_NB <- sapply(1:nrow(ras), function(x){
#   t.test(ras[x, which(colData(NB_GN_dds)$type=="NB")], conf.level = 0.95)$conf.int[1]
# })
# upper_ci_NB <- sapply(1:nrow(ras), function(x){
#   t.test(ras[x, which(colData(NB_GN_dds)$type=="NB")], conf.level = 0.95)$conf.int[2]
# })
# lower_ci_GN <- sapply(1:nrow(ras), function(x){
#   t.test(ras[x, which(colData(NB_GN_dds)$type=="GN")], conf.level = 0.95)$conf.int[1]
# })
# upper_ci_GN <- sapply(1:nrow(ras), function(x){
#   t.test(ras[x, which(colData(NB_GN_dds)$type=="GN")], conf.level = 0.95)$conf.int[2]
# })

pval_df$sd_NB <- sapply(1:nrow(ras_prob_path2), function(x){
  sd(ras_prob_path2[x, which(colData(NB_GN_dds)$type=="NB")])
})
pval_df$sd_GN <- sapply(1:nrow(ras_prob_path2), function(x){
  sd(ras_prob_path2[x, which(colData(NB_GN_dds)$type=="GN")])
})

writexl::write_xlsx(pval_df, "./data/NB_GN_FC_images/NB_GN_ras_prob_path_tbl_FCstatistic.xlsx")
```

```{r fig.width = 10, fig.height=10}
vn.mgraph <- visNetwork::toVisNetworkData(mgraph, idToLabel = FALSE)
vn.mgraph$edges$id <- 1:nrow(vn.mgraph$edges)


vn.mgraph$edges$miriam.kegg.reaction <- unlist(vn.mgraph$edges$miriam.kegg.reaction)
vn.mgraph$edges <- merge(as.data.frame(vn.mgraph$edges), pval_df, by.x = which(colnames(vn.mgraph$edges) =="miriam.kegg.reaction"), by.y = 0, all.x = TRUE)

vn.mgraph$edges$color <- ifelse(vn.mgraph$edges$log2fc < 0, "blue", "red")
vn.mgraph$edges[which(vn.mgraph$edges$padj > 0.05),"color"] <- "grey"
vn.mgraph$edges$value <- abs(vn.mgraph$edges$log2fc)
vn.mgraph$edges$title <- round(vn.mgraph$edges$log2fc,3)

# visNetwork(vn.mgraph$nodes, vn.mgraph$edges) %>%  visEdges(arrows = 'to', smooth =T)
# 
# # compounds of gangliosphingolipid
# comps <- names(keggGet("hsa00604")[[1]]$COMPOUND)
# comps <- gsub("G", "", comps)
# comps[1] <- "C01290"
# 
# vn.mgraph$nodes[which(vn.mgraph$nodes$miriam.kegg.compound == "00127"),"label"] <- "GD1alpha"
# 
# vn.mgraph$nodes <- vn.mgraph$nodes[which(vn.mgraph$nodes$id %in% comps),]
# vn.mgraph$nodes$level <- c(
#   "Lactosylceramide" = 1,
#   "GM1" = 3,
#   "GM2" = 2,
#   "GD3" = 1,
#   "GT2" = 2,
#   "GT3" = 1,
#   "GD1b" = 3,
#   "GT1b" = 4,
#   "GD1a" = 4,
#   "GM1b" = 4,
#   "GM3" = 1,
#   "GD2" = 2,
#   "GA2" = 2,
#   "GA1" = 3,
#   "GT1c" = 3,
#   "GQ1b" = 5,
#   "GQ1balpha" = 5,
#   "GT1aalpha" = 5,
#   "GT1a" = 5,
#   "GD1c" = 5,
#   "GD1a1" = 4)
# # 
# visNetwork(vn.mgraph$nodes, vn.mgraph$edges) %>%
#   visEdges(arrows = 'to', smooth =FALSE, arrowStrikethrough = FALSE, font = list(size=14, align="bottom")) %>%  
#   visNodes(shape = 'box') %>% 
#   visHierarchicalLayout(levelSeparation = NULL, nodeSpacing = 1, edgeMinimization=TRUE,treeSpacing=1) %>%
#   visPhysics(enabled = FALSE)

```

```{r}
mgraph_sub <- create_graph(mgraph)

png("./data/NB_GN_FC_images/NB_GN_graph_ras_prob_path.png", width = 2800, height = 2800, units = "px", pointsize = 20, res="300")
par(omi=c(0,0,0,0), mgp=c(0,0,0),mar=c(0,0,0,0))
plot(mgraph_sub, layout=mylay, vertex.label.color= "black", vertex.label.family= "Arial")
dev.off()
```

```{r}
e_df <- as.data.frame(vn.mgraph$edges)

# ggplot(e_df, aes(x=1, y=miriam.kegg.reaction)) +
#   geom_point(aes(size=-log10(padj), color=log2fc)) +
#   colorspace::scale_color_continuous_diverging(palette = "Blue-Red 3") +
#   theme_classic()
# 
# 
# 
# 
# ggplot(e_df, aes(x=as.character(to), y=as.character(from), fill=log2fc)) +
#   geom_point(aes(size=-log10(padj), shape = 21, colour = "black")) +
#   colorspace::scale_color_continuous_diverging(palette = "Blue-Red 3") +
#   theme_classic() + 
#   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
#   xlab("from node") + ylab("to node")

e_df_lg <- e_df %>%
  tidyr::pivot_longer(colnames(dplyr::select(e_df, c("NB", "GN", "log2fc"))), names_to = "group", values_to = "mean")
e_df_lg$group <- as.factor(e_df_lg$group)
e_df_lg[which(e_df_lg$group %in% c("NB", "GN")),"padj"] <- NA
e_df_lg[which(e_df_lg$group %in% c("log2fc")),"sd_NB"] <- NA
e_df_lg[which(e_df_lg$group %in% c("log2fc")),"sd_GN"] <- NA
e_df_lg[which(e_df_lg$group == "GN"),"sd_NB"] <- NA
e_df_lg[which(e_df_lg$group == "NB"),"sd_GN"] <- NA

e_df_lg <- e_df_lg %>% mutate(sd = coalesce(sd_NB,sd_GN)) 

# ggplot(e_df_lg, aes(x=mean, y=miriam.kegg.reaction, group=group)) + 
#   geom_point(aes(color=-log10(padj))) + 
#   colorspace::scale_color_continuous_sequential(palette = "Reds 3") +
#   facet_wrap(~group, ncol = 3) + 
#   theme_classic() + 
#   theme(panel.grid.major.y = element_line(color = "gray",
#                                           size = 0.5,
#                                           linetype = 1)) + xlab("")


comps <- names(keggGet("hsa00604")[[1]]$COMPOUND)
comps <- gsub("G", "", comps)
comps[1] <- "C01290"
comps <- c(comps, "C01190")

e_df_lg <- e_df_lg[which(e_df_lg$from %in% comps | e_df_lg$to %in% comps),]
e_df_lg$geneSymbol <- sapply(e_df_lg$symbol, function(x) paste0(unique(x), collapse = ' '))
from_label <- c()
for(i in e_df_lg$from){
  from_label <- c(from_label, vn.mgraph$nodes[which(vn.mgraph$nodes$id == i), "label"])
}
e_df_lg$from_label <- from_label
to_label <- c()
for(i in e_df_lg$to){
  to_label <- c(to_label, vn.mgraph$nodes[which(vn.mgraph$nodes$id == i), "label"])
}
e_df_lg$to_label <- to_label

# ggplot(e_df_lg, aes(x=mean, y=paste0(miriam.kegg.reaction, " (", geneSymbol, ")"), group=group)) + 
  
p4 <- ggplot(e_df_lg, aes(x=mean, y=paste0(miriam.kegg.reaction, " (from: ", from_label, " - to: ",to_label, ")"), group=group)) + 
  geom_vline(xintercept = 0, colour="lightgray", linetype="solid") +
  geom_errorbar(aes(xmin=mean - sd,xmax=mean + sd,width=0.2)) +
  geom_point(aes(color=-log10(padj))) +
  colorspace::scale_color_continuous_sequential(palette = "Reds 3") +
  facet_wrap(~group, ncol = 3, scales="free_x") + 
  theme_classic() +
  ylab("") + xlab("") +
  theme(panel.grid.major.y = element_line(color = "lightgray",
                                          size = 0.5,
                                          linetype = 1)) + xlab("")

ggsave(
  "NB_GN_FC_ras_prob_path.png",
  plot = p4,
  device = "png",
  path = "./data/NB_GN_FC_images/",
  scale = 1,
  width = 2800,
  height = 1500,
  units = "px",
  dpi = 300,
  limitsize = FALSE,
  bg = NULL,
)
```


# Multipanel figure

```{r}
# library(multipanelfigure)
# library(grid)
# cols <- 3
# rows <- 4
# figure <- multi_panel_figure(
#   width = 300,
#   columns = cols,
#   height = 450,
#   rows = rows,
#   )
# 
# # figure %<>% fill_panel(p1)
# figure %<>% fill_panel("./data/NB_GNB_FC_images/NB_GNB_FC_ras.png", scaling = "shrink", column = 1:2)
# figure %<>% fill_panel("./data/NB_GNB_FC_images/NB_GNB_graph_ras.png", scaling = "shrink")
# 
# # figure %<>% fill_panel(p2)
# figure %<>% fill_panel("./data/NB_GNB_FC_images/NB_GNB_FC_ras_prob.png", scaling = "shrink", column = 1:2)
# figure %<>% fill_panel("./data/NB_GNB_FC_images/NB_GNB_graph_ras_prob.png", scaling = "shrink")
# 
# # figure %<>% fill_panel(p3)
# figure %<>% fill_panel("./data/NB_GNB_FC_images/NB_GNB_FC_ras_prob_rec.png", scaling = "shrink", column = 1:2)
# figure %<>% fill_panel("./data/NB_GNB_FC_images/NB_GNB_graph_ras_prob_rec.png", scaling = "shrink")
# 
# # figure %<>% fill_panel(p4)
# figure %<>% fill_panel("./data/NB_GNB_FC_images/NB_GNB_FC_ras_prob_path.png", scaling = "shrink", column = 1:2)
# figure %<>% fill_panel("./data/NB_GNB_FC_images/NB_GNB_graph_ras_prob_path.png", scaling = "shrink")
# 
# save_multi_panel_figure(figure, "./data/NB_GNB_FC_images/NB_GNB_FC_image.png", dpi = 600, limitsize=FALSE)
```


